---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

An introduction of the data and a description of the trends/books/items you are choosing to analyze (and why!)

The data set comes from the SPL checkouts from 2017-2023 with monthly updates and still is ongoing. I am choosing to analyze information surrounding number of checkouts, checkout year, material type, and title. Specifically, I am interested in how the frequency of different material formats of books being checked out changes over the years seeing as our society is moving towards increased computerization of everyday things. I am also interested in how the publication date and title impacts the amount of checkouts because I want to know if readers continue to check out books of the same series or are less motivated as the series goes on. Last thing I'm interested in is seeing how the different material types are distributed in terms of the number of checkouts.

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

These will likely be calculated using your DPLYR skills, answering questions such as:

-   What is the average number of checkouts for each item?
-   What is the month or year with the most/least checkouts for a book that you're interested in?
-   What is the month or year with the most/least checkouts for ebooks?
-   How has the number of print book checkouts changed over time?

Feel free to calculate and report values that you find relevant.

### The Dataset

-   Who collected/published the data?\
-   What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
-   How was the data collected or generated?\
-   Why was the data collected?\
-   What, if any, ethical questions do you need to consider when working with this data?\
-   What are possible limitations or problems with this data? (at least 200 words)

The data set was collected and published by the Seattle Public Library. It holds 12 columns of variables that include usage class, checkout type, material type, checkout year, checkout month, number of checkouts, title, ISBN, creator, subjects, publisher, and publication years, of which there are 816,354 rows of observations within the last 6 years spanning from 2017-2023. The data was generated by recording checkout information for each book, audiobook, ebook, etc. every month of each year. As part of the Seattle Open Data Program, the data is collected to accomplish some of the goals such as increasing transparency and promoting research. The SPL face some ethical concerns regarding the homeless population in King County that may want access to the library's services but are not granted them. Library checkouts are not homeless friendly, to sign up for a library card requires entering a permanent address and they cannot afford computers/technologies to checkout digital books online either. Other limitations to the data to consider are some books have more copies than other while some have less therefore some have longer wait times, which could lead to less checkouts. The data also do not tell us if renewals of a previous checkout count as another checkout or just one, or what happens when people lose the books and pay a fine instead.

### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

-   Show more than one, but fewer than \~10 trends
    -   For example, two books, or 5 different mediums
-   You must have clear x and y axis labels
-   The chart needs a clear title 
-   You need a legend for your different line colors and a clear - legend title In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, warning = FALSE}
library('ggplot2')
source('SPL_Chart_1.R')
chart_1

```

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

-   Show more than one, but fewer than \~10 trends
    -   For example, two books, or 5 different mediums
-   You must have clear x and y axis labels
-   The chart needs a clear title 
-   You need a legend for your different line colors and a clear - legend title In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, warning = FALSE}
source('SPL_Chart_2.R')
chart_2

```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

-   You must have clear x and y axis labels
-   The chart needs a clear title 
-   You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE}
source('SPL_Chart_3.R')
chart_3

```
