---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

The data set comes from the SPL checkouts from 2017-2023 with monthly updates and still is ongoing. I am choosing to analyze information surrounding number of checkouts, checkout year, material type, and title. Specifically, I am interested in how the frequency of different material formats of books being checked out changes over the years seeing as our society is moving towards increased computerization of everyday things. I am also interested in how the publication date and title impacts the amount of checkouts because I want to know if readers continue to check out books of the same series or are less motivated as the series goes on. Last thing I'm interested in is seeing how the different material types are distributed in terms of the number of checkouts.

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

These will likely be calculated using your DPLYR skills, answering questions such as:

-   What is the average number of checkouts for each item?
-   What is the month or year with the most/least checkouts for a book that you're interested in?
-   What is the month or year with the most/least checkouts for ebooks?
-   How has the number of print book checkouts changed over time?

Feel free to calculate and report values that you find relevant.

Among some of the values I calculated were average number of library item checkouts for each year, which found that the average checkout value of all the items were kept relatively stable around 23.5 checkouts with a difference of plus or minus 3 and no extremely high or low outliers to be reported. Another thing I evaluated was the total number of checkouts each year for different books within the same series, the results of which make perfect sense. In the years with its maximum checkouts, the first book had the highest number of checkouts with almost 300 then the second book followed with approximately 200, and the last book had the least with less than 150, this possibly shows a reader's lack of interest in continuing the stories. Last, I examined the distribution of a book's material type in all of the checkouts in the last six years, which showed that books in its physical forms were still used most frequently over ebooks and audiobooks. According the Seattle Public Library's database, books have over 100,000 more checkouts than ebooks and the double the amount of audiobooks checked out.

### The Dataset

-   Who collected/published the data?
-   What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
-   How was the data collected or generated?
-   Why was the data collected?
-   What, if any, ethical questions do you need to consider when working with this data?
-   What are possible limitations or problems with this data? (at least 200 words)

The data set was collected and published by the Seattle Public Library. It holds 12 columns of variables that include usage class, checkout type, material type, checkout year, checkout month, number of checkouts, title, ISBN, creator, subjects, publisher, and publication years, of which there are 816,354 rows of observations within the last 6 years spanning from 2017-2023. The data was generated by recording checkout information for each book, audiobook, ebook, etc. every month of each year. As part of the Seattle Open Data Program, the data is collected to accomplish some of the goals such as increasing transparency and promoting research. The SPL face some ethical concerns regarding the homeless population in King County that may want access to the library's services but are not granted them. Library checkouts are not homeless friendly, to sign up for a library card requires entering a permanent address and they cannot afford computers/technologies to checkout digital books online either. Other limitations to the data to consider are some books have more copies than other while some have less therefore some have longer wait times, which could lead to less checkouts. The data also do not tell us if renewals of a previous checkout count as another checkout or just one, or what happens when people lose the books and pay a fine instead.

### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

-   Show more than one, but fewer than \~10 trends
-   For example, two books, or 5 different mediums
-   You must have clear x and y axis labels
-   The chart needs a clear title 
-   You need a legend for your different line colors and a clear - legend title In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, message = FALSE, warning = FALSE, code = readLines("chart1_example.R")}
```


In the first trends over time chart I evaluated average number of checkouts of all the content within the library for each year, the graph of which is separated into these respective years. The graph starts at its lowest point in 2017 and increases steadily into 2018, from 2018 to 2019 it still experiences an increase but a smaller one. It once again increases from 2019 to 2020, which is also the year that peaks in highest average checkouts. The trend declines for the next 2 years and only picks up again in 2023 but that could change seeing as it has only been two months into the current year.

### Second Trends Over Time Chart

```{r, echo = FALSE, message = FALSE, warning = FALSE, code = readLines("chart2_example.R")}
```

In the second trends over time graph, I compared three different books from the same series and author to see if the order would impact how frequently it was checked out as well as answering a question I had about how dedicated readers are to a book series. As you can see, the results were nothing unexpected, they all followed a similar trend pattern reaching their highest checkout numbers in 2020-2021, the first book had the highest number of checkouts, which decreased after each book. Publication date/how old the book is does not seem to matter in terms of its popularity as the oldest book still has many checkouts within the last couple of years. I did however think that the second and the third book would have a similar number of checkouts and the disparity shocked me a little because I have read this books series before and the second book ends somewhat on a cliffhanger and I assumed that readers would be more curious to figure out the issue would be resolved.

### Your Choice

```{r, echo = FALSE, message = FALSE, warning = FALSE, code = readLines("chart3_example.R")}
```

In the third visualization, I chose to make a bar chart in order to see the distribution, whether it was even or skewed a certain direction between the material type of the book and how often it was being checked out. Back then there were only physical books to select from in libraries for borrowing, but now books are offered in other formats that include audiobooks and ebooks, which are both appealing for its convenience of being able to access them virtually anywhere from a phone or tablet. It is because of these reasons that surprise me when my findings show that regular books are still the most popular and most checked out material type considering my own personal experience having grown up with printed books but primarily using ebooks/textbooks whether its for school or leisure. 
according the Seattle Public Library's database with more than 100,000 more checkouts than ebooks and the double the amount of audiobooks checked out.
